Framing Effects in the Wild

### 2022-08-24 paper/Manuscript_Framing.Rmd
Incorporate TM comments (1)

---
title: "Framing Effects in the Wild"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    includes:
      in_header: preamble.tex
# subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "Does news framing affect public opinion? While a large body of experimental work suggests that individual attitudes can be easily altered by emphasizing different aspects of an issue, little is known about how these studies translate into the real world. I exploit a rare natural change of news framing in the major German tabloid newspaper *Bild* to study whether experimental estimates of framing effects translate into the real world. In a first step, I apply supervised transformer models to a novel dataset of 2.5M news articles and provide evidence of a substantial shift in the paper's migration framing following an editorial change. Second, I employ a difference-in-differences design using panel data from over 16,000 individuals in ten waves to assess the precise impact on migration attitudes. The findings show that, despite a major change in migration framing, immigration attitudes and related variables remain unaffected. News framing only seems to be effective within limited opportunity windows during the emergence of political issues."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
---

@@

Does news framing affect public opinion? While a large body of experimental evidence suggests that the emphasis of a certain consideration can substantially alter issue attitudes, we know little about how these findings translate into real-world environments. Prior work has shown that the external validity of such experiments might be seriously complicated by environments of diverse frames [@Chong2013] and actors [@Slothuus2010], as well as the seeking and avoidance of specific content [@Lau2021a; @Taber2006]. Contemporary news environments provide consumers with especially diverse content and maximize their abilities to choose specific outlets and avoid others [@Bennett2008], thereby increasing the gap between the real-world environment and experimental studies. Additionally, the very nature of experiments is prone to over-estimate effects in the real world [@Barabas2010].

Following similar considerations, Donald Kinder called for framing research to extend its toolkit: "[E]xperimental results can always be questioned on their generalizability, and framing effects are no exception. [...] A more balanced reading of frame effects requires methodological diversification, experiments and studies oriented to the world outside" [-@Kinder2007, 157; see also @Busby2019; @Leeper2020]. To the author's knowledge, no study has yet answered this call to provide an estimate of framing effects in the real world.

I address this gap in the literature by exploiting a rare and naturally occuring shift in the migration framing of Germany's largest tabloid newspaper, *Bild*. The case is ideal to assess the everyday influence of issue frames on public opinion, as one outlet substantially changed it's framing while other papers maintained theirs. By employing BERT models for the measurement of crime frames in a corpus of 2.5 million newspaper articles, I am able to clearly identify the treatment - something that is usually left to experimental studies. The fielding of the German Longitudinal Election Study in the same period provides the rare opportunity to precisely assess the effect of this changing framing on readers' migration attitudes in a *difference-in-differences (DiD) design*.

The results indicate that even this strong change in news framing did not affect consumers' immigration attitudes. The estimated effect is statisticially indistinguishable from zero, and significantly smaller than even minimal relevant effect magnitudes. Other variables, such as related issue attitudes, issue importance, and voting intentions are similarly unaffected. Additional evidence suggests that readers do also not react by turning away from the newspaper. Overall, these results fundamentally challenge the external validity of framing effects research and suggest that the influence of individual news outlets on public opinion is rather limited. Even strong changes in news framing do not affect political attitudes or news consumption.

The paper will proceed as follows: in the next section I will briefly introduce the theoretical framework of emphasis framing and derive hypotheses about the impact of news framing on political attitudes. I will then discuss the limited external validity of framing effects studies, before I turn to the research design, including case, measurement, and estimation strategy. The results section will then showcase evidence from difference-in-differences models assessing the impact of changing migration framing on migration attitudes. Lastly, I discuss the wider implications of the findings for the study of framing effects and our understanding of the role of news media in the opinion formation process.

# Emphasis Framing

"Framing is the process by which a communication source, such as a news organisation, defines and constructs a political issue or public controversy" [@Nelson1997Rally]. A "frame" is then the operating unit of this process, defining the issue at hand in a specific way. Different forms of frames have been conceptualised. When I talk about 'frames' here, I refer to *emphasis frames*. These frames emphasise certain topics in relation to an issue, guiding the recipient to think about the issue with those considerations in mind that are promoted by the frame [@Leeper2020, 152]. For example, issues of increasing welfare contributions might be discussed with reference to inequality and providing chances to the poor, or by mentioning that higher welfare benefits might result in higher taxes. Individuals' support of a statement will differ dependent on the frame presented [@sniderman2004structure].

Work using the emphasis framing concept of framing - as opposed to other definitions like equivalence frames [@Kahneman1979prospect], episodic frames [@Iyengar2005], or generic frames [@DeVreese2005] - mostly builds on the expectancy-value model developed by @Ajzen2000. This model suggests that an attitude on a given issue is a function of two things: considerations relating to that object and their respective weights. The evaluation of the issue is subdivided into a number of considerations [a so-called "frame in thought"; @Chong2007a, 105; @Kinder1996], which will carry clear evaluations. For example, to form an opinion about whether immigration to one's country should be restricted, a person might consider the humanitarian conditions in the countries of origin (negative consideration regarding restrictive migration policy), as well as the risk of increased crime (positive consideration; note that the model does not differentiate whether a consideration is grounded in objective reality). Individuals weigh each of these considerations to form an overall opinion on the subject.

Figure \ref{fig:cogStor} visualizes this logic. A given issue is associated with an attitude $y$. The issue is related to a number of different considerations, each of which carries an associated evaluation $x_i$, e.g. that crime is bad. Each of these considerations is more or less associated with the issue. Based on the strength of this association $w_i$, a given consideration's evaluation is more or less reflected in the issue evaluation $y$, which is equal to the weighted sum of considerations.

```{=tex}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{vis/CognitiveStorage.png}
    \caption{The cognitive evaluation of an issue $y$ is determined by the strength of association $w_i$ with other concepts with existing evaluations $x_i$.}
    \label{fig:cogStor}
\end{figure}
```
How does this model incorporate news content and other forms of political communication? News content can emphasize different considerations of an issue [so called "frames in communication" @Kinder1996]. This strengthens the cognitive association of the issue at hand with the raised consideration. As a result, this specific consideration factors more prominently in an individuals' attitude formation process: "media frames influence opinions by stressing specific values, facts, or other considerations, endowing them with greater apparent relevance to the issue than they might appear to have under an alternative frame" [@Nelson1997Rally, 569].

Applying this to the issue of migration policy and the consideration of criminal migrants, one can imagine a news consumer confronted with an increased volume of news about crimes committed by migrants. The value-expectancy model predicts that the consumer forms a cognitive definition of migration that is increasingly associated with the issue of crime, meaning this consideration features more prominently when forming an opinion about the appropriate level of immigration policy. As a result, their opinion about migration policies should become more restrictive.

***Association hypothesis*****: When an individual is exposed to a specific consideration about an issue, their association of the issue with this consideration should be strengthened.**

***Evaluation hypothesis*****: When an individual is exposed to a specific positive/negative consideration about an issue, their attitude about the overall issue should become more positive/negative.**

@@

Research on framing has shown for decades that citizens' attitudes and their issue associations are highly reactive to minor changes in the emphasis of specific considerations. However, it remains unclear how well these experimental findings translate into the real world. Especially when it comes to the exposure to frames in communication, translating framing effects from experimental studies is only possible under strong assumptions: While framing studies usually simply present a news article or other forms of political content to respondents, news content in the real world competes with other news and entertainment content to reach its audience. Additionally, changes in news framing itself might affect how likely readers are to consume news content: readers might be less likely to consume frames challenging their existing attitudes and reject news outlets containing such frames (and vice versa seek out attitudinally congruent information). This idea of 'selective exposure/avoidance' [@Bennett2008], or a a 'confirmation bias' [@Taber2006] brings with it a major endogeneity problem making purely correlational evidence of framing effects rather uninformative.

Even if we assume that the consumer receives the content, e.g. buys a newspaper, visits its website, or is exposed otherwise, it is not clear whether the content is actually consumed. E.g., a reader might well buy a newspaper but skip an article after reading the headline because the topic is not of interest or the frame is at odds with their existing attitude [@Lau2021a]. As most framing studies do not offer choice of content to respondents, engagement remains an assumption in experimental studies. Additionally, even when consuming the content, motivated reasoners presented with a plethora of news content about an issue will spend less time and cognitive energy on counter-attitudinal arguments than those supporting their viewpoint, lending more importance to considerations raised by amenable content [@Taber2006]. Experimental studies presenting only one frame per issue cannot take this bias into account.

These assumptions of exposure, consumption, and equal engagement in classic framing studies already represent a high hurdle for the translation of experimental estimates of framing effects into the real world. However the differences do not end there. Consumption might result in better or worse comprehension in the real world because respondents might pay more or less attention to real news content compared to experimental treatments [@Barabas2010]. Even after understanding the frame, there might be differences in how easily this translates into changing cognitive associations, given that those exposed might take real-world content more seriously.

To summarize, framing might have very different consequences in real-world environments than their experimental emulations. When we take the experimental evidence as an approximation of the effectiveness of framing in the real world, we make strong assumptions about citizens' exposure to, consumption of, and engagement with the frames in news content. Experiments can only incorporate few characteristics of the real world at a time. In order to test the external validity of framing effect studies, it is necessary to study observational data.

@@

Several scholars have arrived at this conclusion and called for the assessment of framing effects beyond experimental studies [@Busby2019; @Kinder2007; @Leeper2020], yet the call has rarely been answered. While a large body of media effects research exists, it has not paid specific interest to framing, instead assessing the impact of changing news exposure broadly in observational designs [see e.g. @Foos2022; @Grossman2022 @Levendusky2022; @Spirig2020] and field experiments [@Bail2018; @Broockman2022; @Guess2021].

The precious small body of existing research on framing effects is often riddled with endogeneity concerns or compound treatments. For example, @Jerit2008 investigates the effects of framing on public opinion compared to addressing the opponent's issue. She finds that reframing is less effective than engagement with the opposite parties' frames, a finding that suggests that framing is not very effective. However, this study does not provide an estimate of a causal effect, instead restricting itself to the assessment of correlational evidence. This is especially problematic regarding framing effects, as news content can affect consumers' inclination to consume news (see preceding section), and political actors might react to public opinion. Others show with similar correlational designs that only specific frames seem to drive issue importance [@Mclaren2018], but do not assess effects on attitudes.

To my knowledge, there are only two observational studies assessing the causal effect of exposure to specific content on political attitudes. @Slothuus2010b takes advantage of a sudden shift in a political parties' framing of retirement benefits, and shows that this changed framing has substantial impact on party supporters' attitudes about the issue. However, it is unclear how these findings translate to news framing effects, as it is impossible to distinguish the effect of partisan cue-taking from the persuasive effect of the frame itself. The only study which assesses the precise impact of changing news content on political attitudes comes from @Djourelova2020. She exploits a ban of the term "illegal migrant" in the *Associated Press* and shows that this language has modest effects on migration attitudes. While the study itself reveals impressively how language can shape political attitudes, it constitutes a highly specific case and a rather minor change in the framing of migration. While news stories might lack the term "illegal migrant", they might still extensively cover crimes committed by migrants and thus associate migration with crime.

Summarized, we lack an understanding of how framing affects political opinions in the real world. While experimental evidence suggests rather strong effects, it is unclear how these findings translate to the outside world. Observational evidence on media effects is abundant but has not studied precisely what the impact of framing is. The present study therefor exploits a naturally occurring change in the news framing of a particular issue *within* a specific outlet to isolate the effects of news framing on political opinions.

@@

In order to test whether changing news content - rather than a shift in the consumed news outlet - affected political attitudes, I exploit an editorial change in the largest German newspaper, *Bild*. As I will show, this editorial change resulted in a substantial shift in the papers' migration framing and is hence ideal to assess the effect of within-newspaper content shifts. The fielding of a large-scale panel survey in the same period allow to assess the impact of news content on migration attitudes.

@@

In December 2016, Axel Springer, the company owning Bild, announces that Diekmann will leave the company. Less than a week after, it becomes public that prosecutors investigate Diekmann for sexual harassment. While the company claimed that this was not the reason for Diekmann's departure, the process might have been sped up as a result of the investigations\footnote{According to a former editor of the newspaper I talked to, rumors in the company at the time support this interpretation of events.} [@Spiegel2017Diekmann]. The migration coverage in *Bild* severely changed when Diekmann left the newspaper in February 2017 and Julian Reichelt took over the position as editor-in-chief. The paper increasingly started painting a picture of refugees as criminals on the tabloid's front page [@Niggemeyer2018Bild; @DLF2018Reichelt]. Headlines like "I killed 40 people and want asylum" increasingly dominated the migration coverage of the newspaper. In a speech months before he took over the position as chief editor, Reichelt outlined his motivation for such a shift in migration coverage: "I can assure you: nothing has hurt [*Bild*] economically as much as our clear, humane, empathetic stance in the refugee crisis" \footnote{Reichelt in a speech at Deutschlandfunk's conference "Formate des Politischen 2016": \url{https://vimeo.com/190347766}}. He also argued that the German media shows too little tolerance towards anti-immigration attitudes and therefor excludes those holding such views from the political discourse.

This presents a unique case to study the effects of media content on public opinion. The shift in content can be clearly identified. The change of editor itself was not discussed in the media, which means any opinion changes among readers are the result of the changing content [@Spirig2020]. Additionally, the timing of the editorial change is potentially exogenous, given the pressure from allegations against Diekmann. In summary, *this case is as close as possible to a field experiment in which a specific consideration of migration coverage is amplified in the coverage of one newspaper, but not others*.

@@

To identify whether an article is about migration, i employ supervised machine learning. As migration content is rare, and training data needs to be somewhat balanced, articles cannot be selected randomly for annotation[^1]. Instead, I first construct a dictionary of terms related to migration. I use dictionary extension based on German GloVe word-embeddings\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from articles with no migration terms, 100 from the quarter with most migration terms and 100 from the remainder in between) for each newspaper for a total of 1,800 papers. Then, a coder hand-coded these articles, assessing whether their main topic is related to migration.

@@

Figure \@ref(fig:salience) shows the share of migration articles in each newspaper. Clear peaks in attention are visible surrounding the refugee movements of summer 2015, as well as following the large-scale sexual assaults on new years eve in Cologne 2016[^2], and a conflict over asylum policy between the conservative sister parties CDU and CSU, which almost lead to the breakdown of the governing coalition in 2018. Newspapers seem to be event-driven and follow largely parallel trends, albeit at different levels, with the centre-left \textit{SZ} discussing migration most and the tabloid \textit{Bild} least. Attention decreases in mid-2016, but remains high compared to pre-2015 levels. Surrounding the fielding period of the survey waves (gray vertical lines), most newspapers show similar levels of attention to the issue between 2.5% and 6% of all articles, with the exception of \textit{SZ}, which moves between 5% and 12.5%.

[^2]: Germany had a large debate about refugees and crime after the new year's eve 2015/16 in Cologne. According to a report of the ministry of interior in Northrhine-Westphalia, hundreds of women were sexually assaulted and robbed by men they perceived to be of "north-african/arabic" descent (see report of the ministry of interior in Northrhine-Westphalia: \\url{[https://web.archive.org/web/20160116104020/http://m.mik.nrw.de/fileadmin/user_upload/Redakteure/Dokumente/Themen_und_Aufgaben/Schutz_und_Sicherheit/160111ssia/160111berppkoeln.pdf}](https://web.archive.org/web/20160116104020/http://m.mik.nrw.de/fileadmin/user_upload/Redakteure/Dokumente/Themen_und_Aufgaben/Schutz_und_Sicherheit/160111ssia/160111berppkoeln.pdf%7D%7D))

To identify migration coverage containing the crime consideration and thereby associating migration with crime, I first fine tune a BERT classifier for German language in a similar fashion as above, this time with coverage about crime as the outcome of interest. Using a training set of 1400 articles, a validation set of 200 and a test set of 200 articles, the classifier reaches a very good performance with an F1 score of 85.7%. I then apply this classifier to classify the approximately 90,000 articles about migration identified by the first classifier. About 7% of all migration coverage consists of articles about some form of crime\footnote{See appendix 1.2 for more information about the classifier, training process and data.}. The share of migration content devoted to news about crime within a given time frame allows me to assess how prominent the consideration of crime as a result of migration was at a specific point in time. In order to give an example of the content, table \@ref(tab:exampleArticles) displays the headlines of the five migration articles within *Bild* most likely discussing crime according to the classifier. All articles refer to crimes of migrant groups, with two referring to violence in refugee homes, two discussing police raids regarding asylum fraud, and one discussing a police report about new years eve in cologne.

>>>>

---
title: "Framing Effects in the Wild"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    includes:
      in_header: preamble.tex
# subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "Does news framing affect public opinion? While a large body of experimental work suggests that individual attitudes can be easily altered by emphasizing different aspects of an issue, little is known about how these studies translate into the real world. I exploit a rare natural change of news framing in the major German tabloid newspaper *Bild* to study whether experimental estimates of framing effects translate into the real world. In a first step, I apply supervised transformer models to a novel dataset of 2.5M news articles and provide evidence of a substantial shift in the paper's migration framing following an editorial change. Second, I employ a difference-in-differences design using panel data from over 16,000 individuals in ten waves to assess the precise impact on migration attitudes. The findings show that, despite a major change in migration framing, immigration attitudes and related variables remain unaffected. News framing seems to affect attitudes only within limited windows of opportunity early in the politicization of political issues."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
nocite: |
  @CologneReport2016
---

@@

Does news framing affect public opinion? While a large body of experimental evidence suggests that the emphasis of a certain consideration can substantially alter issue attitudes, we know little about the translation of these findings into real-world environments. Prior work has shown that the external validity of such experiments might be seriously complicated by environments of diverse frames [@Chong2013] and actors [@Slothuus2010], as well as the seeking and avoidance of specific content [@Lau2021a; @Taber2006]. Especially on contested issues, citizens might be more inclined to resist persuasion [@Arceneaux2009]. Contemporary news environments provide consumers with especially diverse content and maximize their abilities to choose specific outlets and avoid others [@Bennett2008], thereby increasing the gap between the real-world environment and experimental studies. Additionally, the very nature of experiments is prone to over-estimate effects in the real world [@Barabas2010]. This is the reason why, fifteen years ago, Donald Kinder called for framing research to extend its toolkit: "[E]xperimental results can always be questioned on their generalizability, and framing effects are no exception. [...] A more balanced reading of frame effects requires methodological diversification, experiments and studies oriented to the world outside" [-@Kinder2007, 157; see also @Busby2019; @Leeper2020].

I provide the - to my knowledge - first observational estimate of framing effects by exploiting a rare shift in the migration framing of Germany's largest tabloid newspaper, *Bild*. The case is ideal to assess the everyday influence of issue frames on public opinion, as one outlet substantially changed its framing while other papers maintained theirs. By employing BERT models for the measurement of crime frames in a corpus of 2.5 million newspaper articles, I am able to clearly identify the treatment - something that is usually left to experimental studies. The fielding of the German Longitudinal Election Study in the same period provides the rare opportunity to precisely assess the effect of this changing framing on readers' migration attitudes in a *difference-in-differences (DiD) design*.

[might want to mention frames in mind and frames in communication distinction here and how it informs research design (follow Rune & Martin Bisgaard)]

The results indicate that even a strong increase in the use of crime frames did not affect consumers' immigration attitudes. The estimated effect is statistically indistinguishable from zero and significantly smaller than even minimal effect magnitudes. Related issue attitudes, voting behaviour, and issue importance are similarly unaffected, as is the association of immigration attitudes with the frame provided. Additional evidence implies that immigration attitudes had crystallized at the time of the shift, such that individuals were immune to changing framing.

Overall, these results fundamentally challenge the external validity of framing effects research and suggest that the influence of individual news outlets on public opinion is more limited than previously thought. Even a strong change in news framing did not affect political attitudes. Instead, it seems that timing is a crucial factor in the effectiveness of news frames.

# Framing Effects on Public Opinion

"Framing is the process by which a communication source, such as a news organisation, defines and constructs a political issue or public controversy" [@Nelson1997Rally]. Different forms of frames have been conceptualized. When I talk about 'frames' here, I refer to *emphasis frames*. These frames emphasize certain topics in relation to an issue, guiding the recipient to think about the issue with those considerations in mind that are promoted by the frame [@Leeper2020, 152]. For example, issues of increasing welfare contributions might be discussed with reference to inequality and providing chances to the poor, or by mentioning that higher welfare benefits might result in higher taxes. Individuals' support of a statement will differ dependent on the frame presented [@sniderman2004structure]. Many other forms of framing exist, such as equivalence frames [@Kahneman1979prospect], episodic frames [@Iyengar2005], or generic frames [@DeVreese2005]. The focus on emphasis frames in this study is owed to the large body of literature and generalized theory regarding their effect on attitudes about any issue (as opposed to generic and episodic frames), as well as their prevalence in the news media (which is not the case for equivalence frames).

Work using the emphasis framing concept of framing mostly builds on the expectancy-value model developed by @Ajzen2000. This model suggests that an attitude on a given issue is a function of two things: the value of considerations relating to that issue and the strength of their association with the issue. The evaluation of the issue is subdivided into a number of considerations [a "frame in mind"; @Chong2007a, 105; @Nelson1996], which will carry clear evaluations. For example, to form an opinion about whether immigration to one's country should be restricted, a person might consider the humanitarian conditions in the countries of origin (negative consideration regarding restrictive migration policy), as well as the risk of increased crime (positive consideration; note that the model does not differentiate whether a consideration is grounded in objective reality). Individuals weigh each of these considerations to form an overall opinion on the subject.

Figure \ref{fig:cogStor} visualizes this logic. A given issue (bottom square) carries an attitude $y$. The issue is associated with a number of different considerations (top circles), each of which carries an evaluation $x_i$, e.g. crime is viewed very negatively. Based on the strength of the association $w_i$ of a given consideration with the issue, a given consideration's evaluation is more or less reflected in the issue evaluation $y$, which is equal to the weighted sum of considerations. For example, if a given individual associates immigration strongly with an increase in criminal activity, welfare fraud, but only very weakly with humanitarian considerations, cultural enrichment, liberty, or economic growth, they will form a rather negative attitude about the issue, as the former factor more prominently in their reasoning about migration.

```{=tex}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{vis/CognitiveStorage.png}
    \caption{The cognitive evaluation of an issue $y$ is determined by the strength of association $w_i$ with other concepts with existing evaluations $x_i$.}
    \label{fig:cogStor}
\end{figure}
```

How does this model incorporate news content and other forms of political communication? News content can emphasize different considerations of an issue [so called "frames in communication" @Nelson1996]. By being exposed to the consideration, an individual's cognitive association of the issue at hand with the raised consideration strengthens (or in more technical terms: the weight of the consideration $w_i$ increases). As a result, this specific consideration factors more prominently in the formulation of an overall attitude: "media frames influence opinions by stressing specific values, facts, or other considerations, endowing them with greater apparent relevance to the issue than they might appear to have under an alternative frame" [@Nelson1997Rally, 569].

Applying this to the issue of migration policy and the consideration of criminal migrants, one can imagine a news consumer confronted with an increased volume of news about crimes committed by migrants. The value-expectancy model predicts that 1) the consumer forms a cognitive definition of migration that is increasingly associated with the issue of crime, meaning this consideration features more prominently when forming an opinion about the appropriate level of immigration; 2) as a result, their opinion about migration policies should become more restrictive, as the negative consideration of crime is weighted more heavily in the evaluation of the migration issue.

***Association hypothesis*****: When an individual is exposed more to a specific consideration about an issue, their association of the issue with this consideration should be strengthened.**

***Evaluation hypothesis*****: When an individual is exposed more to a specific positive/negative consideration about an issue, their attitude about the overall issue should become more positive/negative.**

@@

Research on framing has shown for decades that citizens' attitudes and their issue associations are highly reactive to minor changes in the emphasis of specific considerations [@Chong2007; @Chong2010; @Druckman2012Bias; @Kinder1990; @Nelson1996; @Nelson1997; @sniderman2004structure; for reviews see @Busby2019; @Druckman2001; @Leeper2020]. However, it remains unclear how well these experimental findings translate into the real world. First, taking experimental effects as estimates of the real-world impact of framing means making strong assumptions about citizens' exposure to frames: while framing studies usually simply present a news article or other forms of political content to respondents, news content in the real world competes with other news and entertainment content to reach its audience. Changes in news framing itself might affect how likely readers are to consume news content: readers might be less likely to consume frames challenging their existing attitudes and reject news outlets containing such frames (and vice versa seek out attitudinally congruent information). This idea of 'selective exposure/avoidance' [@Bennett2008], or a 'confirmation bias' [@Taber2006] brings with it a major endogeneity problem making purely correlational evidence of framing effects rather uninformative.

Second, even if one assumes that the consumers receive the content, (e.g. buy a newspaper, visit a website, ...), it is not clear whether the content is actually consumed. For example, a reader might well buy a newspaper but skip an article after reading the headline because the topic is not of interest or the frame is at odds with their existing attitude [@Lau2021a]. As most framing studies do not offer choice of content to respondents, engagement remains an assumption when taking experimental studies as evidence of framing effects in the real world (for an exception see @Druckman2012Bias). Additionally, information from different sources, such as TV, social media, blogs, newspapers, ..., might 'crowd out' the framing of a single outlet: By consuming content from many sources, individuals might give less attention to a single outlet, thereby severely limiting the overall impact of that outlet's coverage.

Third, even when consuming the content, cognitive engagement with it might strongly differ in the real world compared to experimental contexts. Motivated reasoners presented with a plethora of news content about an issue will spend less time and cognitive energy on counter-attitudinal arguments than those supporting their viewpoint, lending more importance to considerations raised by amenable content [@Taber2006]. Experimental studies presenting only one frame per issue cannot take this bias into account. Consumption might also result in better or worse comprehension in the real world because respondents might pay more or less attention to real news content compared to experimental treatments [@Barabas2010]. Furthermore, there might be differences in how easily this translates into changing cognitive associations, given that those exposed might take real-world content more seriously. Additionally, the relative timing at which a frame is received, relative to other frames and the salience of the issue, affects the way in which individuals react to framing. Those citizens who received a different frame earlier might already have formed a crystallized opinion about the issue and thus reject subsequent attempts of persuasion [@Bechtel2015; @Druckman2012Bias; @Lecheler2009b].

To summarize, there are many reasons to assume that framing might have very different consequences in real-world environments than in their experimental emulations, and the above list is by no means comprehensive. When we take the experimental evidence as an approximation of the effectiveness of framing in the real world, we make many strong assumptions about citizens' exposure to, consumption of, and engagement with the frames in news content. Experiments can only incorporate few characteristics of the real world at a time. In order to test the external validity of framing effect studies, it is necessary to study observational data.

@@

Despite repeated calls for the assessment of framing effects beyond experimental studies [@Busby2019; @Kinder2007; @Leeper2020], there is little research addressing this concern. While a large body of observational media effects research exists, it has not paid specific interest to framing, instead assessing the impact of changing news exposure broadly [@Bail2018; @Broockman2022; @Foos2022; @Grossman2022; @Guess2021; @Levendusky2022; @Spirig2020].

The precious small body of existing observational research on framing effects is often riddled with endogeneity issues. For example, @Jerit2008 investigates the effects of news framing on public opinion. She finds that framing is less effective in changing public opinion than engagement with opposing considerations, a finding that suggests that framing is rather ineffective. However, her study restricts itself to the assessment of correlational evidence. This is especially problematic regarding framing effects, as news content can affect consumers' inclination to consume news (see preceding section), and political actors might react to public opinion, thereby creating endogeneity issues. Others show with similar designs that only specific frames seem to drive issue importance [@Mclaren2018], but do not assess effects on attitudes.

To my knowledge, there are only two observational studies assessing the causal effect of exposure to specific content on political attitudes. @Slothuus2010b takes advantage of a sudden shift in a political parties' framing of retirement benefits, and shows that this changed framing has substantial impact on party supporters' attitudes about the issue. However, it is unclear how these findings translate to news framing effects. Given that - in this case - a positional change coincided with the changing framing of the issue and parties' special role in citizens' information process, it is impossible to distinguish the effect of partisan cue-taking from the persuasive effect of the frame itself.

The only study which assesses the precise impact of changing news content on political attitudes comes from @Djourelova2020. She exploits a ban of the term "illegal migrant" in the *Associated Press* and shows that this language has modest effects on migration attitudes. While the study itself reveals impressively how language can shape political attitudes, it constitutes a highly specific case and a rather minor change in the framing of migration. To gain a deeper understanding of how specifically framing affects immigration attitudes, it is necessary to assess broader changes in the representation of the immigration issue.

In sum, we lack an understanding of how framing affects political opinions in the real world. While experimental evidence suggests rather strong effects, it is unclear how these findings translate to the outside world. Observational evidence on media effects is abundant but has not precisely studied the impact of framing. The present study therefor exploits a naturally occurring change in the news framing of a particular issue *within* a specific outlet to isolate the effects of news framing on political opinions.

@@

In order to test whether changing news content affected political attitudes, I exploit an editorial change in the largest German newspaper, *Bild*. As I will show, this editorial change resulted in a substantial shift in the papers' migration framing and is hence ideal to assess the effect of within-newspaper content shifts. The fielding of a large-scale panel survey in the same period allow to assess the impact of news content on migration attitudes.

@@

In December 2016, Axel Springer, the company owning Bild, announced that Diekmann will leave the company. Less than a week after, it became public that prosecutors investigate Diekmann for sexual harassment. While the company claimed that this was not the reason for Diekmann's departure, the process might have been sped up as a result of the investigations\footnote{According to a former editor of the newspaper I talked to, rumors in the company at the time support this interpretation of events.} [@Spiegel2017Diekmann]. The migration coverage in *Bild* severely changed when Diekmann left the newspaper in February 2017 and Julian Reichelt took over the position as editor-in-chief. The paper increasingly started painting a picture of refugees as criminals on the tabloid's front page [@Niggemeyer2018Bild; @DLF2018Reichelt]. Headlines like "I killed 40 people and want asylum" increasingly dominated the migration coverage of the newspaper. In a speech months before he took over the position as chief editor, Reichelt outlined his motivation for such a shift in migration coverage: "I can assure you: nothing has hurt [*Bild*] economically as much as our clear, humane, empathetic stance in the refugee crisis" \footnote{Reichelt in a speech at Deutschlandfunk's conference "Formate des Politischen 2016": \url{https://vimeo.com/190347766}}. While he expressed his support for the decision to take this stance at the time, he argued that the German media shows too little tolerance towards anti-immigration attitudes and therefore excludes those holding such views from political discourse.

This case presents a unique case to study the effects of media content on public opinion. As will be shown below, a shift in migration coverage can be clearly identified. The change of editor itself was not discussed in the media, which means any opinion changes among readers are the result of the changing content [@Spirig2020]. Additionally, the timing of the editorial change is potentially exogenous, given the pressure from allegations against Diekmann. In summary, *this case is as close as possible to a field experiment in which a specific consideration of migration coverage is amplified in the coverage of one newspaper, but not others*.

[Outline research design, measurement of frames in communication and thought, following RUne and Martin Bisgaard]

@@

To identify whether an article is about migration, I employ supervised machine learning. As migration content is rare, and training data needs to be somewhat balanced, articles cannot be selected randomly for annotation[^1]. Instead, I first construct a dictionary of terms related to migration. I use dictionary extension based on German GloVe word-embeddings\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from articles with no migration terms, 100 from the quarter with most migration terms and 100 from the remainder in between) for each newspaper for a total of 1,800 papers. Then, a coder hand-coded these articles, assessing whether their main topic is related to migration.

@@

Figure \@ref(fig:salience) shows the share of migration articles in each newspaper. Clear peaks in attention are visible surrounding the refugee movements of summer 2015, as well as following the large-scale sexual assaults on new years eve in Cologne 2016\footnote{Germany had a large debate about refugees and crime after the new year's eve 2015/16 in Cologne. According to a report of the ministry of interior in Northrhine-Westphalia, hundreds of women were sexually assaulted and robbed by men they perceived to be of "north-african/arabic" descent (Ministerium für Inneres und Kommunales des Landes Nordrhein-Westfalen 2016).}, and a conflict over asylum policy between the conservative sister parties CDU and CSU, which almost led to the breakdown of the governing coalition in 2018. Newspapers seem to be event-driven and follow largely parallel trends, albeit at different levels, with the centre-left \textit{SZ} discussing migration most and the tabloid \textit{Bild} least. Attention decreases in mid-2016, but remains high compared to pre-2015 levels. Surrounding the fielding period of the survey waves (gray vertical lines), most newspapers show similar levels of attention to the issue between 2.5% and 6% of all articles, with the exception of \textit{SZ}, which moves between 5% and 12.5%.

To identify migration coverage containing the crime consideration and thereby associating migration with crime, I first fine tune a BERT classifier for German language in a similar fashion as above, this time with coverage about crime as the outcome of interest. Using a training set of 1400 articles, a validation set of 200 and a test set of 200 articles, the classifier reaches a very good performance with an F1 score of 85.7%. I then apply this classifer to identify the approximately 90,000 articles about migration identified by the first classifier. About 7% of all migration coverage consists of articles about some form of crime\footnote{See appendix 1.2 for more information about the classifier, training process and data.}. The share of migration content devoted to news about crime allows me to assess the prominence of the consideration of crime within migration coverage. As an example of the content, table \@ref(tab:exampleArticles) displays the headlines of the five migration articles within *Bild* which were identified to discuss crime with the highest likelihood. All articles refer to crimes of migrant groups, with two referring to violence in refugee homes, two discussing police raids regarding asylum fraud, and one discussing a police report about new years eve in cologne.

```{=tex}
\begin{table}
\centering
\begin{tabular}{|p{12cm}|}
\hline
Final new years eve report of Cologne police: male groups identified \\ \hline
Police searching for Northern-african offender: asylum-raid in Dortmund! \\ \hline
Search in refugee home - mega-raid addressing asylum fraud \\ \hline
Two groups attacking each other - again bloody conflict in refugee shelter \\ \hline
Cabin fever! Violence! Mass brawls! This is behind the increase in refugee clashes \\ \hline
\end{tabular}
\caption{Headlines of five migration articles most likely to concern crime, according to BERT crime classifier. Translation by author.}
\label{tab:exampleArticles}
\end{table}
```

The left panel of figure \@ref(fig:treatment) shows the share of migration coverage devoted to crime news (solid, red line), once for Bild and once for all other newspapers (dashed, black line). As we can see, *Bild*'s attention to crime news in its migration coverage substantially increased briefly following the takeover of Julian Reichelt, while it focused less on the issue preceding the editorial change. The trends run largely parallel to each other preceding the treatment, with the major exception being the few months following the events at new year's eve in Cologne 2015/16\footnote{Note that the delayed increase in crime coverage and the low levels of attention to crime in *Bild*'s migration coverage in 2016 are partially the result of a changing number of articles being published. This can be traced back to a changing share of content from the German news agency *dpa*, which provides services to quickly distribute news content similar to the Associated Press. Estimates excluding content from *dpa* indicate a change in attention to crime of similar size, as can be seen in appendix 1.}. The right-hand panel shows difference-in-difference estimates of the crime emphasis in migration coverage of one newspaper compared to all others, before and after the editorial change in Bild. One can see that *Bild* was the only newspaper which substantially changed its emphasis of the frame, while other newspapers maintained their levels of attention.
